Collaborative Visual Analysis on the Web with RCloud

### 2015-03-31 discussion.tex
working

On top of this, RCloud allows or even encourages hackers to fork experiments
to try new ideas quickly. So far we have not done much more to address the
problem of having a lot of bits and pieces of code and data lying around,
though we have provided a framework in which it should be easier to find them
and to apply algorithms and metadata to organize them.
\stephen{I think we need to defend that a bit more.}

@@

Our back-end environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome.
Information security in RCloud should be improved, but more
sophisticated approaches (such as information flow analysis,
and query languages that ensure differential privacy) are
active research topics, and many difficult problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
new approaches to information security in visual analytics.
\gordon{This sounds like we think we are helping by pushing the
system to the breaking point, is that the idea?}

@@

The interface is also perceived to be slower, both for computer and for
human. First, there is a slower reaction to commands being run, because
RCloud necessarily saves a cell to its repository before executing it. It
may be possible to safely write to the repository behind execution of the
same code. Second, and perhaps more detrimental, the interface requires
mouse interaction and switching between the keyboard and mouse, which are
known to be slower. \gordon{we don't need to back that up, right? i hate
  using passive voice.} Shortcut keys can be devised if people are willing
to learn them, but it would be better to reduce the number of such
operations needed.

Other complaints mostly focus on layout and can probably be addressed with
some iteration and customization features.

{\bf Discovery is hard.}
While we have provided tools for searching and curating notebooks, the
search function does not help for finding methodology (and does not protect
searchers from erroneous notebooks), and the notebook tree starts to get
unwieldy with hundreds of users with hundreds of notebooks. In addition,
our choice to retain notebooks even when they have been deleted has
exacerbated the situation, since obsolete and broken notebooks stay around
forever.

One thing to try is tagging and filtering the tree on tags, so that users
can mark their own notebooks or those of others to make them easier to
find. However, this relies on users making a conscious effort, and it may be
better to rely more on passive metrics such as how often a notebook has
been run.

The system also needs to help users curate their notebooks. If a notebook
has been deleted or superseded by one of its forks, a user who opens it
should be warned. And operations on multiple notebooks and folders of
notebooks should be supported to make curation easier.

{\bf Collaboration is hard.}
While forking provides a simple way to continue someone else's work, working
in teams is a lot more complicated. The forking feature is popular but it
poses a problem, too, as dozens of versions of a notebook start proliferating.

@@

More general solutions would be serious engineering endeavors. Even
GitHub's generous web interface does not support merging where there are
conflicts. Actual sharing of notebooks would probably require moving toward
git's notion of decentralized code repositories, because GitHub gists do
not support shared ownership.

>>>>

On top of this, RCloud allows or even encourages hackers to fork experiments
to try new ideas quickly. So far we have not done much more to address the
problem of having a lot of bits and pieces of code and data lying around,
though we have provided a framework in which it should be easier to find them
and to apply algorithms and metadata to organize them.

@@

Our back-end environment assumes a high degree of trust between users.
Access control for information security is delegated to the host
operating system and web server. In practice, most operating
systems and web servers rely either on coarse permission models,
which tend to be ineffective, or on detailed access control
lists, which tend to be cumbersome.
Information security in RCloud should be improved. More
sophisticated approaches, such as information flow analysis,
and query languages that ensure differential privacy, are
active research topics, and many problems remain
\cite{Moore:2011:SAF}. We are hopeful that having a richer
environment for collaboration and information sharing will encourage
new approaches to information security in visual analytics.
%\gordon{This sounds like we think we are helping by pushing the
%system to the breaking point, is that the idea?}

@@

The interface is also perceived to be slower, both for computer and for
human. First, there is higher latency when commands run, because
RCloud saves a cell to a repository before executing it. It
may be possible to safely write to the repository behind execution of the
same code. Second, and perhaps more detrimental, the interface requires
switching between the keyboard and mouse, which can be cumbersome.
%\gordon{we don't need to back that up, right? i hate using passive voice.}
Shortcut keys can defined as a workaround, but it would be better
to reduce the need for them.

Other complaints mostly focus on layout, and can probably be addressed by
offering better customization, and designs such as tabbed views.

{\bf Discovery is hard.}
While we provided tools for searching and curating notebooks, the
search function does not help for finding methodology (and does not protect
searchers from erroneous notebooks), and the notebook tree starts to get
unwieldy with hundreds of users and many hundreds of notebooks. In addition,
our choice to retain notebooks even when they have been deleted 
exacerbates the situation, since obsolete and broken notebooks stay around
forever.

One thing to try is tagging and filtering on tags, so that users
can mark their own notebooks or those of others to make them easier to
find. However, this relies on users making a conscious effort, which
is not always reliable. It seems better if possible to lean more on
passive metrics, such as how often a notebook has been run.

The system also needs to help users to curate notebooks. If a notebook
has been deleted or superseded by one of its forks, a user who opens it
should be warned. And operations on multiple notebooks and folders of
notebooks should be supported to make curation easier.

{\bf Collaboration is hard.}
While forking provides a simple way to continue someone else's work, working
in teams is a lot more complicated. The forking feature is popular but it
poses a problem, too, as dozens of versions of a notebook proliferate.

@@

A more general solutions would be a serious engineering endeavor. Even
GitHub's admirable web interface does not support merging where there are
conflicts. Unrestricted sharing of notebooks would probably require moving
toward git's notion of decentralized code repositories, because GitHub gists
do not support shared ownership.
